|

### 2020-08-08 pre-analysis-plan/freire-skarbek-lynching.Rmd
update pap, address comments egap and livia

---
title:  |
        | Pre-Analysis Plan:
        | Casting the First Stone: Understanding Attitudes towards Lynching in Brazil
author:
  - Danilo Freire[^freire]
  - David Skarbek[^skarbek]
date: \today
abstract: "Why do citizens resort to extrajudicial violence? While vigilantism has significantly decreased in advanced democracies, lynchings continue to be common practice in many developing nations. In this pre-analysis plan, we propose three survey experiments to investigate (i) who are the potential lynching victims, (ii) how citizens justify mob violence, and (iii) whether information provision can reduce support for vigilantism. First, we will run a conjoint experiment to assess which victim characteristics are associated with higher likelihood of lynching. Then we will show respondents a short news article about a real lynching episode and ask them whether they perceive the motivations of the lynch mob as legitimate. Lastly, we test whether providing information about legal punishment for lynching perpetrators, human rights guarantees, and the risk of retribution makes respondents less likely to endorse extralegal violence. We will run the experiments in Brazil, a country which has seen a sharp rise in vigilante attacks and currently experiences one lynching attempt per day. Our survey will be conducted online and include 1,000 participants."
abstractspacing: double
keywords: Brazil, crime, extralegal violence, lynching, vigilantism
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: template.latex
---

@@

@@

Why do citizens resort to extralegal punishment?[^definitions] Although mob
violence has decreased in advanced democracies, vigilantism remains widespread
in the developing world. In particular, Latin America has been severely affected
by a recent increase in lynchings. Motivated by growing numbers of drug-related
crime and by police ineffectiveness, citizen violence has surged in the region
[@mallen2014vigilantes]. Brazil provides a telling example. From 2011 to 2015,
the country registered about 2,500 lynching episodes, and in 2015 alone 173
people were killed by angry mobs, an average of one execution every two days
[@barbara2015vigilantes; @zechmeister2014political]. According to José Martins,
who has studied lynchings in Brazil for more than 30 years, those figures are
not only the highest in the country's history, but also the highest in the world
[@pearson2018latam]. Victims are generally accused of petty theft, and many are
beaten to death for stealing mobile phones, cooking utensils, or pairs of
sandals [@barbara2015vigilantes]. Lynching perpetrators are often young men, but
they also include female teenagers, elderly women, and even members of local
police forces [@oliveira2016mob].

@@

Recent research suggests that support for vigilantism in Latin America is
associated with exposure to violence [@garcia2019anger], untrustworthy law
enforcement [@zizumbo2017community], perceived insecurity [@ceobanu2011crime;
@godoy2004justice], and social inequality [@arias2010violent]. However, approval
for extrajudicial violence raises important questions about the relationship
between states and societies in the region. First, while vigilantism provides
social order to vulnerable populations, it also undermines the legitimacy of
state institutions [@schuberth2013challenging]. Lynching episodes reinforce the
idea that civilians should not wait for legal proceedings and that social
control is better exerted by the people themselves [@black1983crime]. Second,
victim characteristics greatly impact the likelihood and severity of lynchings.
In Brazil, a person is rarely lynched when they commit a crime against males,
and females that commit the same crimes as males also receive less
punishment.[^fapesp] Moreover, there is no significant social difference between
victims and perpetrators, nor black Brazilians are particularly targeted by
lynch mobs [@oliveira2016mob]. These dynamics pose a puzzle to existing
explanations for mob violence support and stand in sharp contrast with the
American experience [@smaangs2016doing; @tolnay1995festival; @wood2011lynching].

@@

In this study, we endeavour to answer the following questions: 1) which victim
characteristics increase individual support for extralegal punishment?; 2) how
do individuals justify their approval for vigilantism?; and 3) can information
provision reduce support for mob violence? We design a set of three experiments
to tackle these issues. Our first experiment consists of a conjoint analysis
where we present respondents with different profiles of lynching victims and ask
them which individual they believe deserved the punishment. In our second
experiment, we test how ineffective policing, slow criminal justice systems,
and demand for hasher punishment for criminals affect individual propensity to
support mob violence. Finally, we assess whether citizens who support
extrajudicial killings change their opinions when informed about the legal
consequences of lynchings, constitutional human rights, or the risk of
_vendettas_ associated with mob violence.

@@

We start our analysis with a choice-based conjoint experimental design. Each
respondent will be shown ten pairs of lynching victims. Each profile will
consist of five attributes: 1) crime perpetrator; 2) race of the crime
perpetrator; 3) offense; 4) victim of the motivating crime; 5) lynching
perpetrators. The levels in each attribute are available in table
\ref{tab:categories} below. They will be randomised in the experiment. Each
respondent will read a short description of the survey experiment and we will
ask them to choose the profile in which lynching seems less justified. We phrase
the experiment in negative terms (who should _not_ be lynched) to protect
subjects from making the uncomfortable choice of actively selecting a lynching
victim.

\begin{table}[htpb]
\small
\begin{center}
\caption{\textbf{Attributes and Levels}}
\label{tab:categories}
\begin{tabular}{l !{\vrule width 1pt}p{9cm}}
\Xhline{2\arrayrulewidth}
\textbf{Attribute} & \multicolumn{1}{c}{\textbf{Levels}} \\
\Xhline{2\arrayrulewidth}
\footnotesize
Crime perpetrator & Adult male; adult female; male teenager; female teenager; elderly man; elderly woman \\
\\
Race & Black; white; native Brazilian; Asian \\
\\
Offense & Picks the pocket; steals the car; molests; rapes; murders \\
\\
Crime victim & Adult male; adult female; male child; female child; male teenager; female teenager; elderly man; elderly woman\\
\\
Lynching perpetrators & Bystanders; acquaintances; the family of the victim;
police\\
\Xhline{2\arrayrulewidth}
\end{tabular}
\end{center}
\end{table}

@@

Respondents are also required to rate each profile on a scale from 1
to 7, where 1 indicates they definitely believe lynching was not justified
and 7 indicates they definitely believe lynching was justified. We take this
measure as a robustness check for our main experimental estimate.

Before each profile pair, respondents will read the following
message:

* Lynching is often used as social punishment in Brazil. We are interested in
  knowing more about how Brazilians see these episodes. In the next ten
  questions, please read the description of two possible lynching victims in
  Brazil and indicate in which case you believe the punishment is _less_
  justified.[^portuguese]

[^portuguese]: The text will be displayed in Portuguese. The translation is as
follows: Linchamento é por vezes usado como punição social no Brasil. Estamos
interessados em saber mais sobre como os brasileiros vêem estes episódios. Nas
próximas cinco questões, por favor, leia a descrição de duas possíveis vítimas
de linchamento no Brasil e indique em quais delas você acredita que a punição
é _menos_ justificada.

@@

We invite respondents from all regions of Brazil to participate in our
experiment. Qualtrics will recruit 1,000 adult Brazilian citizens to take part
in our survey experiment.

@@

We will estimate our models with the `cregg` package [@leeper2018cregg] for the
R statistical language [@rstats2019]. We will report marginal means instead of
average marginal conditional effects (AMCE) of attitudes towards lynchings.
@leeper2018subgroup show that AMCEs can be misleading in subgroup analysis as
model results are sensitive to the choice of reference categories in
interactions. In contrast, marginal means provide a clear description of
quantities of interest, in our case preferences towards lynching, while allowing
for easy comparisons between groups of respondents. Their interpretation is also
straightforward: a 50% marginal means estimate represents that respondents are
indifferent when a given attribute appears vis-à-vis other attributes. When the
coefficient is lower than 50%, respondents dislike profiles with that attribute.
Conversely, when the point estimate is higher than 50%, respondents prefer
profiles containing a given attribute. We will cluster standard errors by
respondent.

@@

We expect that, on average, respondents will select profiles with the following
characteristics:

1) Crime perpetrator: male teenager
2) Race: blacks
3) Offense: murder
4) Victim: female child
5) Lynching perpetrators: the family of the victim

We also expect that punishment will be lower for females of any age, whites
and Asians, pick pocketers, and if the victim of the motivating crime is male.
Additionally, we expect that respondents see lynchings as less legitimate if
they are carried out by bystanders and the police forces. We have no prior
beliefs about the size of each coefficient.

@@

We will also run subgroup comparisons in our sample. We hypothesise that
attitudes towards lynching vary according to age, gender, race, level of
education, political ideology, and previous victimisation. In particular, we
hypothesise that:

@@

* Respondents who were victimised in the last year are more likely to support
  lynchings by bystanders and the police.

@@

In our second experiment, we will measure how respondents justify their
support for extralegal violence. We test the effect of three explanations for
lynching in Brazil: 1) police ineffectiveness; 2) slow criminal justice; 3)
demand for harsher punishment for criminals. Below, we discuss these hypotheses
in further detail.

@@

Another aspect that is often suggested in the literature is that vigilantism
rises where citizens have little faith in courts. Most commonly, this is due to
delays in the criminal justice administration, which makes criminals stay at
large for many years. In Brazil, the penal code allows the accused to appeal
each decision several times, so it can take decades before a criminal case is
closed [@sousa2005utilizaccao]. In this respect, citizens do not believe that,
even if the criminal is put to trial, he/she will be punished in a timely
matter. Note that although the police is part of the criminal justice system, we
separate the two institutions in our experiment. This is because respondents may
not associate the two, mainly when it comes to legal procedures against
lawbreakers.

Lastly, we evaluate whether respondents believe that the legal punishment
assigned to criminals is not proportional to the severity of the crimes. In
particular, we hope to gauge the demand for iron-fisted criminal justice in
Brazil. Although this treatment arm is related to the previous ones, it
questions not the efficiency of the institutions, but their legitimacy. We
believe that the demand for harsher laws is on the rise in Brazil, and this may
be a major reason why citizens justify their support for mob violence.
In fact, Brazilians are often vocal about their preference for repressive legal
punishment. After having his son murdered by a local thug, a bar owner told the
Wall Street Journal that "_even if he had been put behind bars for 100 years it
wouldn't have been enough to pay for all his crimes_" [@pearson2018latam]. We
hypothesise that this feeling is widespread in the population.

The experiment consists of three treatment conditions and one control group.
Respondents will read a short excerpt of a news article describing a real
lynching case, which will be slightly edited so that respondents have no prior
knowledge of the crime.[^jornal] The vignette for the control group includes no
information about the reasons behind the lynching. We will ask respondents
to show their level of support for mob violence using a 0-100 slider, where 0
means no support and 100 means full support. Respondents in each of the three
treatment arms will read the same piece, but with one additional sentence
explaining the motivation of the lynch mob. The vignettes are as follows:

[^jornal]: The original article is available at the following address:
<http://folhadeibitinga.com.br/materia/5543/suspeito-de-estuprar-menina-e-linchado-e-morto-em-jau>.
Access: June 2020.

* *Control group*: A young man between 20 and 25 years old was lynched and
  killed on the 1\textsuperscript{st} of June. He was accused of raping an
  11-year-old girl in São Paulo. He was clubbed to death by about 20 residents
  who heard the girl screaming after she was taken to an empty lot and forced
  to have sex with him at gunpoint.[^control-exp02]

[^control-exp02]: In Portuguese: Um jovem entre 20 e 25 anos foi linchado e
morto no dia primeiro de junho. Ele foi acusado de estuprar uma menina de 11
anos em São Paulo. Ele foi atacado e morto à pauladas por cerca de 20 moradores
que ouviram a menina gritar após ela ter sido levada para um terreno baldio e
forcada a ter relações sexuais com ele sob ameaça de morte.

* *Treatment 01 - Police ineffectiveness*: A young man between 20 and 25
  years old was lynched and killed on the 1\textsuperscript{st} of June. He was
  accused of raping an 11-year-old girl in São Paulo. He was clubbed to death
  by about 20 residents who heard the girl screaming after she was taken to an
  empty lot and forced to have sex with him at gunpoint. **One of the
  residents who took part in the lynching said they did so because "the police
  never patrols the area"**.[^treat01-exp02]

[^treat01-exp02]: In Portuguese: Um jovem entre 20 e 25 anos foi linchado e
morto no dia primeiro de junho. Ele foi acusado de estuprar uma menina de 11
anos em São Paulo. Ele foi atacado e morto à pauladas por cerca de 20 moradores
que ouviram a menina gritar após ela ter sido levada para um terreno baldio e
forçada a ter relações sexuais com ele sob ameaça de morte. **Um dos moradores
envolvidos no linchamento disse que eles fizeram isso porque "a polícia nunca
patrulha o local".**

* *Treatment 02 - Criminal justice ineffectiveness*: A young man between 20
  and 25 years old was lynched and killed on the 1\textsuperscript{st} of June.
  He was accused of raping an 11-year-old girl in São Paulo. He was clubbed to
  death by about 20 residents who heard the girl screaming after she was taken
  to an empty lot and forced to have sex with him at gunpoint. **One of the
  residents who took part in the lynching said they did so because "the
  judicial system is too slow"**.[^treat02-exp02]

[^treat02-exp02]: In Portuguese: Um jovem entre 20 e 25 anos foi linchado e
morto no dia primeiro de junho. Ele foi acusado de estuprar uma menina de 11
anos em São Paulo. Ele foi atacado e morto à pauladas por cerca de 20 moradores
que ouviram a menina gritar após ela ter sido levada para um terreno baldio e
forçada a ter relações sexuais com ele sob ameaça de morte. **Um dos moradores
envolvidos no linchamento disse que eles fizeram isso porque "a justiça é muito
lenta"**.

* *Treatment 03 - Demand for harsher legal punishment*: A young man between 20
  and 25 years old was lynched and killed on the 1\textsuperscript{st} of June.
  He was accused of raping an 11-year-old girl in São Paulo. He was clubbed to
  death by about 20 residents who heard the girl screaming after she was taken
  to an empty lot and forced to have sex with him at gunpoint. **One of the
  residents who took part in the lynching said they did so because "the
  judicial punishment is not harsh enough"**.[^treat03-exp02]

[^treat03-exp02]: In Portuguese: Um jovem entre 20 e 25 anos foi linchado e
morto no dia primeiro de junho. Ele foi acusado de estuprar uma menina de 11
anos em São Paulo. Ele foi atacado e morto à pauladas por cerca de 20 moradores
que ouviram a menina gritar após ela ter sido levada para um terreno baldio e
forcada a ter relações sexuais com ele sob ameaça de morte. **Um dos moradores
envolvidos no linchamento disse que eles fizeram isso porque "a punição da
justiça não é dura o suficiente"**.

@@

1) Older citizens
2) White
3) Low levels of education
4) Right-wingers
5) Had been previously victimised

@@

Our last experiment aims to measure the effect of information provision on
attitudes about lynching. More specifically, we will test whether reminding
respondents about the legal and social consequences of mob violence reduces
the subjects' level of support for such practice.

The experiment has three treatment conditions plus a control group. In all of
them we will present respondents with a short statement affirming that some
Brazilians support vigilantism under certain conditions. The control group will
be asked to rate their agreement with the statement with a 0-100 slider where
100 indicates full agreement with the statement and 0 indicates no agreement.
Respondents will be asked to use 0 to 49 if they disagree, 50 if they neither
agree nor disagree, and 50-100 if they agree with the sentence.

@@

\noindent _H1_: Providing information about the criminal consequences of
lynchings reduces individual support for vigilantism.

@@

\noindent _H2_: Reminding respondents that torture violates human rights does
not decrease their support for vigilantism.

@@

\noindent _H3_: Information about _vendettas_ decrease individual support for
vigilantism.

@@

These three hypotheses will allow us to test how respondents react to different
consequences of extralegal legal and, hopefully, also help policy-makers design
more effective campaigns against vigilantism in Brazil.

>>>>

---
title:  |
        | Pre-Analysis Plan:
        | Casting the First Stone: Understanding Attitudes towards Lynching in Brazil[^thanks]
author:
  - Danilo Freire[^freire]
  - David Skarbek[^skarbek]
date: \today
abstract: "Why do citizens support extrajudicial violence? While endorsement for vigilantism has significantly decreased in advanced democracies, lynchings remain tolerated throughout the developing world. In this pre-analysis plan, we propose three survey experiments to investigate (i) which criminal profiles citizens prefer for extralegal punishment, (ii) how individuals justify mob violence, and (iii) whether information provision reduces support for vigilantism. First, we run a conjoint experiment to identify criminal characteristics that are associated with higher likelihood of lynching. In our second experiment, we show respondents a short article about a real lynching episode and ask them whether they see the mob's motivations as legitimate. Lastly, we test whether providing information about the legal consequences of lynching, human rights guarantees, and the risk of retribution makes respondents less likely to endorse extralegal violence. We will run the experiments in Brazil, a country which has seen a sharp rise in vigilante attacks and currently experiences one lynching attempt per day. Our survey will be conducted online and include 2,000 participants."
abstractspacing: double
keywords: Brazil, crime, extralegal violence, lynching, vigilantism
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: template.latex
---

@@

[^thanks]: We thank Rosario Aguilar, Miriam Golden, Umberto Mignozzetti, Irfan Nooruddin, Catarina Roman, and Alex Sacco for their valuable comments.

@@

Why do citizens endorse extralegal punishment?[^definitions] Although mob
violence has decreased in advanced democracies, vigilantism remains widespread
in the developing world. In particular, Latin America has been severely affected
by a recent increase in lynchings. Motivated by growing numbers of drug-related
crime and by police ineffectiveness, citizen violence has surged in the region
[@mallen2014vigilantes]. Brazil provides a telling example. From 2011 to 2015,
the country registered about 2,500 lynching episodes, and in 2015 alone 173
people were killed by angry mobs, an average of one execution every two days
[@zechmeister2014political]. According to José Martins, who has studied
lynchings in Brazil for more than 30 years, those figures are not only the
highest in the country's history, but also the highest in the world
[@pearson2018latam]. Victims are generally accused of petty theft, and many are
beaten to death for stealing mobile phones, cooking utensils, or pairs of
sandals [@barbara2015vigilantes]. Lynching perpetrators are often young men,
but they also include female teenagers, elderly women, and even members of
local police forces [@moura2017linchamentos; @oliveira2016mob].

@@

Recent research suggests that support for vigilantism in Latin America is
associated with exposure to violence [@garcia2019anger], untrustworthy law
enforcement [@zizumbo2017community], perceived insecurity [@ceobanu2011crime;
@godoy2004justice], and social inequality [@arias2010violent]. However, approval
for extrajudicial violence raises important questions about the relationship
between states and societies in the region [@nivette2016institutional]. First,
while vigilantism provides social order to vulnerable populations, it also
undermines the legitimacy of state institutions [@schuberth2013challenging].
Lynching episodes reinforce the idea that civilians should not wait for legal
proceedings and that social control is better exerted by the people themselves
[@black1983crime]. Second, victim characteristics greatly impact the likelihood
and severity of lynchings. In Brazil, a person is rarely lynched when they
commit a crime against males, and females that commit the same crimes as males
also receive less punishment.[^fapesp] Moreover, there is no significant social
difference between victims and perpetrators, nor black Brazilians are
particularly targeted by lynch mobs [@oliveira2016mob]. These dynamics pose a
puzzle to existing explanations for mob violence support and stand in sharp
contrast with the American experience [@smaangs2016doing; @wood2011lynching].

@@

In this study, we endeavour to answer the following questions: 1) which victim
characteristics increase individual support for extralegal punishment?; 2) how
do individuals justify their approval for vigilantism?; and 3) can information
provision reduce support for mob violence? We design a set of three experiments
to tackle these issues. Our first experiment consists of a conjoint analysis
in which we show respondents different profiles of lynching victims and ask them
who they believe deserve the punishment. In our second experiment, we test how
ineffective policing, slow criminal justice systems, and demand for hasher legal
punishment affect individual propensity to support mob violence. Finally, we
assess whether citizens become less supportive of vigilantism when they are
informed about the legal consequences of lynchings, constitutional human rights,
or the risk of _vendettas_ associated with mob violence. While in this
pre-analysis plan we present the three experiments as described above,
their order will be randomised in our survey to mitigate potential order bias.

@@

We start our analysis with a choice-based conjoint experimental design. We will
present five pairs of criminal profiles to respondents. Each profile consists
of seven attributes: 1) gender of the crime perpetrator; 2) age of the crime
perpetrator; 3) race of the crime perpetrator; 4) offense; 5) gender of the victim
of the motivating crime; 6) age of the victim of the motivating crime; 7)
lynching perpetrators. The attributes and levels are displayed in table
\ref{tab:categories} below.

\begin{table}[htpb]
\small
\begin{center}
\caption{\textbf{Attributes and Levels}}
\label{tab:categories}
\begin{tabular}{l !{\vrule width 1pt}p{9cm}}
\Xhline{2\arrayrulewidth}
\textbf{Attribute} & \multicolumn{1}{c}{\textbf{Levels}} \\
\Xhline{2\arrayrulewidth}
\small
Gender of crime perpetrator & Male; female \\ [4pt]
Age of crime perpetrator & Teenager; adult; elderly \\ [4pt]
Race of crime perpetrator & Black; white; native Brazilian; Asian \\ [4pt]
Offense & Picks the pocket; steals the car; molests; rapes; murders \\ [4pt]
Gender of crime victim & Male; female\\ [4pt]
Age of crime victim & Child; teenager; adult; elderly\\ [4pt]
Lynching perpetrators & Bystanders; neighbours; family of the victim; gangs; police \\
\Xhline{2\arrayrulewidth}
\end{tabular}
\end{center}
\end{table}

@@

Respondents have to indicate which profile they prefer for extrajudicial
punishment. They are are also required to rate each profile on a scale from 1 to
7, where 1 indicates they definitely believe lynching was not justified and 7
indicates they definitely believe lynching was justified. We take this measure
as a robustness check for our main experimental estimate.

Respondents will read the following prompt before they start the experiment:

* Lynchings are often used as social punishment in Brazil. Lynchings are
  cases in which three or more people physically attack or execute a suspected
  criminal in public. We are interested in knowing more about how Brazilians see
  these episodes. In the next five questions, please read the description of two
  possible lynching victims in Brazil and indicate in which case you believe the
  punishment is more justified. Even if you are not entirely sure, please select
  one of the cases.[^portuguese]

[^portuguese]: We will conduct the survey experiments in Portuguese. The
translation for the first experiment is as follows: Linchamentos são às vezes
usados como punição social no Brasil. Linchamentos são casos nos quais três ou
mais pessoas agridem fisicamente ou executam em público um suspeito de um
crime. Estamos interessados em saber mais sobre como os brasileiros vêem estes
episódios. Nas próximas cinco questões, por favor, leia a descrição de duas
possíveis vítimas de linchamento no Brasil e indique em quais delas você
acredita que a punição é mais justificada. Mesmo que você não tenha certeza,
por favor, escolha um dos casos.

@@

We invite respondents from all regions of Brazil to participate in our
experiment. Qualtrics will recruit 2,000 adult Brazilian citizens to take part
in our survey experiment.

@@

We will estimate our models with the `cregg` package [@leeper2018cregg] for the
R statistical language [@rstats2019]. We will report marginal means and 
average marginal component effects (AMCEs) in our main analysis.
However, @leeper2018subgroup show that AMCEs can be misleading in subgroup
comparisons as model results are sensitive to the choice of reference categories in
interactions. In contrast, marginal means provide a clear description of
quantities of interest, in our case preferences towards lynching, while allowing
for easy comparisons between groups of respondents. So while we will include
both AMCEs and marginal means estimators in our article, we will conduct our
subgroup analysis using only the latter. We will cluster standard errors by
respondent.

@@

We expect that, on average, respondents are more likely to select profiles with
the following attributes levels:

1) Gender of crime perpetrator: male 
2) Age of crime perpetrator: teenager
3) Race of crime perpetrator: blacks
4) Offense: murder
5) Gender of victim: female 
6) Age of victim: child
7) Lynching perpetrators: the family of the victim

We also expect that support for extralegal punishment will be lower for females
of any age, whites and Asians, pick pocketers, and if the victim of the
motivating crime is male. Additionally, we expect that respondents see lynchings
as less legitimate if they are carried out by bystanders, gangs, or the police
forces. We have no prior beliefs about the size of each coefficient.

@@

We also intend to run subgroup analysis in our sample. We believe that
preferences towards lynching may vary according to age, gender, race, level of
education, political ideology, and previous victimisation. In particular, we
hypothesise that:

@@

* Respondents who were victimised in the last year are more likely to support
  lynchings by neighbours and the police.

@@

In our second experiment, we analyse how respondents justify their preferences
for extralegal violence. We assess the impact of three factors that have been
cited as major drivers of vigilante justice in Brazil: 1) police
ineffectiveness; 2) slow criminal justice; 3) demand for harsher punishment for
criminals. Below, we discuss them in further detail.

@@

Another determinant of support for vigilante justice that is suggested in the
literature is the lack of trust in the justice system [@godoy2004justice]. This
is often due to long criminal proceedings, which cause significant anxiety for
the victims. In Brazil, the penal code allows the accused to appeal each
decision several times, so it can take decades before a criminal case is closed
[@sousa2005utilizaccao]. In this respect, citizens do not believe that, even if
the criminal is put to trial, he/she will be punished in a timely matter. Note
that although the police is part of the criminal justice system, we separate the
two institutions in our experiment.

Lastly, we evaluate whether respondents think that the legal punishment
assigned to criminals is not proportional to the severity of their crimes. In
particular, we hope to gauge the demand for iron-fisted criminal justice in
Brazil. Although this treatment arm is related to the previous ones, it
addresses not the efficiency of the institutions, but their legitimacy
[@nivette2016institutional]. We believe that the demand for harsher laws is on
the rise in Brazil, and this may be a major reason why citizens justify their
support for mob violence. In fact, Brazilians are often vocal about their
preference for repressive legal punishment. In a recent article in
_The Wall Street Journal_, a bar owner justified the lynching of the local thug
who killed his son by saying that "_even if he had been put behind bars for 100
years it wouldn't have been enough to pay for all his crimes_"
[@pearson2018latam]. We hypothesise that many Brazilians also share this view.

The experiment consists of three treatment conditions and one control group.
Respondents will read an excerpt of a news article describing a real
lynching case. We have slightly edited the original text so that respondents
have no prior knowledge of the crime.[^jornal] The vignette for the control
group includes no information about the reasons behind the lynching. We will ask
respondents to show their level of support for mob violence using a 0-100
slider, where 0 means no support and 100 means full support. Respondents in each
of the three treatment arms will read the same piece, but with one additional
sentence explaining the motivations behind the lynching. The vignettes are as
follows:

[^jornal]: The original article is available at the following address:
<https://jr.jor.br/2020/05/01/homem-e-linchado-na-vila-progresso>.
Access: August 2020.

* *Control group*: A man was lynched last Friday in Jundiaí, São Paulo.
  According to the neighbours, he tried to break into a house but was
  immobilised and beaten by members of the community.[^control-exp02]

[^control-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade.

* *Treatment 01 - Police ineffectiveness*: A man was lynched last Friday in
  Jundiaí, São Paulo. According to the neighbours, he tried to break into a
  house but was immobilised and beaten by members of the community. **One of
  the residents who took part in the lynching said they had beaten the suspect
  because "the police never patrols the area"**.[^treat01-exp02]

[^treat01-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a polícia
nunca patrulha o local".**

* *Treatment 02 - Criminal justice ineffectiveness*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community.
  **One of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial system is too slow"**.[^treat02-exp02]

[^treat02-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a justiça
é muito lenta"**.

* *Treatment 03 - Demand for harsher legal punishment*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community. **One
  of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial punishment is not harsh
  enough"**.[^treat03-exp02]

[^treat03-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a punição
da justiça não é dura o suficiente"**.

@@

1) Older citizens
2) White
3) Low levels of education
4) Right-wingers
5) Individuals who had been previously victimised

@@

Our last experiment aims to measure the effect of information provision on
attitudes about lynching. More specifically, we will test whether reminding
respondents about the legal and social consequences of vigilante justice reduces
the subjects' level of support for such practice.

The experiment has three treatment conditions and a control group. In all of
them we will present respondents with a short statement affirming that some
Brazilians support vigilantism under certain conditions. The control group will
be asked to rate their agreement with the statement with a 0-100 slider where
100 indicates full agreement with the statement and 0 indicates no agreement.
Respondents will be asked to use 0 to 49 if they disagree, 50 if they neither
agree nor disagree, and 50-100 if they agree with the sentence.

@@

\noindent _H1_: Individuals who receive information about the criminal
consequences of lynching show lower support for vigilantism that those in the
control group.

@@

\noindent _H2_: Reminding respondents that torture violates human rights has no
impact on their support for vigilantism.

@@

\noindent _H3_: Information about _vendettas_ triggered by lynchings makes
individuals less likely to support vigilantism in comparison with respondents in
the control group. The effect of that information is larger than the other
two treatments.

@@

These three hypotheses allow us to test how respondents react to different
consequences of extralegal violence. Hopefully, the experiment may also help
policy-makers design more effective campaigns against vigilantism in Brazil.
